<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Functional Programming with mchof}
-->

Functional Programming with mchof
=================================

## 1 Introduction

## 2 Installation

You have presumably installed mchof if you are reading this vignette; if not
you can simply type
````
install.packages('mchof')
```
Into the console. It is also possible to install the 
development version of mchof by using `install_github('mchof', 'rgrannell1)`, 
though this is not recommended; the development version is usually unstable. 
Updates will be released roughly every month, for the first few releases.

## 3 Functional Programming Basics




## 4 Using `mchof`

### 4.1 Parallel Options

### 4.2 Multiple Argument Functions

### 4.3 NA Handling

### 4.4 Strings as Function Names

Functions like mcFold and mcReduce take binary functions such as plus or multiply.
It would be frustrating to have to write `get('+')` every time you wanted to call
an infix operator, so all mchof functions also accept a function name
in place of an actual function. Here's how you would sum the vector 
1,...,10 with mcReduce and mcFold respectively.

```{r}
mcReduce('+', seq_len(10))

mcFold('+', 0, seq_len(10), list(mc.cores = 2))
```





```{r css}
options(markdown.HTML.stylesheet = system.file('misc','vignette.css', package='knitr'))
```

